\documentclass[12pt]{labbook}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage{float}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{listings}
\usepackage{color}
\usepackage{array}
\usepackage{times}
\usepackage{amssymb, amsmath}
\usepackage{fancyhdr}
\usepackage[]{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{courier}

\usepackage{marginnote}%provides ability to put notes in the margin using \marginnote{} command

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\textwidth=16.5cm

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\title{Notes for Undergraduate Research Work}
\author{Alan Dixon}

\begin{document}

\section{June 6, 2016 Notes}

\begin{itemize}
    \item Discovered and fixed two bugs involving an uninitialized selection vector and Mphi values
    \item Said bug fixes resolved the issue of the program crashing around iteration 2500
      
    \item Despite the fixed bugs resolving that issue, log likelihood values still plummet to values eventually reaching 10e-23 which causes the not a number error, causing the MCMC algorithm to halt.
\marginnote{mikeg: This is confusing given the previous statement. 
  It sounds like you didn't really resolve the underlying issue.}  
    \item Due to Newton failing to execute my job request, I had to run the R script on my laptop so I was only able to run it once so more testing is needed but for the meantime it appears it takes until around iteration 42000 for the program to crash.
      \marginnote{why is newton failing to execute your job requests? 
        How and when will you solve this issue?}
    \end{itemize}

TODO:\\
    Work on psuedocode\\
    Remove redundancies\\
    Determine if Sphi prior is being used\\
    Improve varialbe names and documentation\\
    Document the R scripts\\
    
\section{June 7, 2016}

\begin{itemize}
    \item Removed exit statements from FONSEParameter.cpp, PANSEParameter.cpp, and RFPParameter.cpp in order to eliminate warnings when running the check in R
    \item Downloaded all of the necessary files to my Gauley account.
    \item Running FONSE with simulated data and b = 0.001 led to "nan" being reached much more quickly than on my laptop, around iteration 5500.
      \marginnote{mikeg: is this based on one run on each machine?
        It's more useful to be more quantitative and descriptive} 
    \item Said issue may be a cause of R not working correctly on Gauley as compared to my laptop. 
      Multiple errors and missing packages were encountered. 
      These errors made it impossible to check any plots.
\item{mikeg: If it strange you're running into such errors since gauley is used to develop the code base.
    Are they really errors or, since you execute code, warnings?
    What are the error codes?  
    When discussing errors, it is good to always include some output using the 'quote' or similar environment in \LaTeX.
    Do other people run into these same errors?
  }
    \item Two new Items for the TODO list: Continue troubleshooting R on Gauley and alter Jeremy's .sge scripts so that they no longer email him when I run a job on Newton.
\end{itemize}

\section{June 8, 2016}

\begin{itemize}
    \item Changed the .sge scripts for Newton so that Jeremy no longer receives an email upon the completion of a job.
    \item Changed a function in FONSEModel.cpp to match it's corresponding function in ROCModel.cpp. In particular, I changed a long series of divisions to on division and a series of multiplications because multiplication is a faster operation.
      \marginnote{mikeg: include the function name. 
        Any information on speed up?
      }
    \item Fixed a typo in the documentation of ROCModel.cpp
    \item Worked out some kink experienced while running R scripts on Gauley. In particular, finally made it so that the plots actually showed up and worked. 
    \item Formulated the hypothesis that the reason we are experiencing a drop in log likelihood comes from a bug involving either deltaM, deltaOmega, or phi after observing the trace plots.
\marginnote{Include plots in your notes, especially if you refer to them, and explain more clearly why you think its behavior indicates this.
}
\end{itemize}
\end{document}